---
title: "Understanding Michaelis–Menten Kinetics"
author: "Martin De Kauwe"
output: html_document
---

# Introduction

Michaelis–Menten kinetics describes how the rate of an enzyme-catalyzed reaction 
depends on substrate concentration, capturing this relationship with two key 
parameters: the maximum reaction rate (V<sub>max</sub>) and the substrate 
concentration at which the reaction rate is half of V<sub>max</sub> 
(K<sub>m</sub>). This framework provides a useful analogy for understanding how 
Rubisco activity is influenced by CO<sub>2</sub> availability in C<sub>3</sub>
photosynthesis.

```{r}
library(ggplot2)
```

We are going to start by describing our model as a function. A model is a 
simplified mathematical description of a process - in this case, how an 
enzyme's reaction rate depends on substrate concentration. A function in R is a 
reusable block of code that takes inputs (called arguments), performs a 
calculation, and returns an output. Writing the model as a function makes it 
easy to apply the same formula to many different values or to reuse in other 
code

```{r}
michaelis_menten <- function(S, Vmax, Km) {
  # Model is after Leonor Michaelis and Maud Menten
  
  # S is the substrate
  # Vmax is the maximum reaction rate (i.e., when the enzyme is saturated)
  # Km is the Michaelis constant, the substrate concentration at half Vmax
  V <- Vmax * S / (Km + S)

  return (V)
}
```

Let's now run our model...
```{r}

# Parameters 
Vmax <- 1.5   # μmol/min/mg enzyme
Km <- 0.05    # mM 

# Create an array of substrate concentrations (S, mM)
substrate_conc <- seq(0, 0.5, by=0.01)

# check what this looks like
print(substrate_conc)

# Compute reaction velocities
velocities <- michaelis_menten(substrate_conc, Vmax, Km)

# Create a data frame to make it easy to store the data and plot
df <- data.frame(
  substrate = substrate_conc,
  velocity = velocities
)

```

And make a plot to see how our model works
```{r}
ggplot(df, aes(x=substrate, y=velocity)) +
  geom_line(color = "seagreen", linewidth=1.5) +
  labs(
    x = "Substrate concentration [S] (mM)",
    y = "Reaction rate, V (μmol/min/mg enzyme)"
  ) +
  theme_classic()
  

```

## Questions

1. What happens to the reaction rate when the substrate concentration is 
low vs high?
2. If we change V<sub>max</sub> how does it affect our plot? Hint: Change 
V<sub>max</sub> and plot the old line and the new model prediction.
3. How does changing K<sub>m</sub> affect the affinity for the substrate? Hint: 
Change K<sub>m</sub> and plot the old line and the new model prediction.

## Summary

By exploring the Michaelis–Menten model, we have learned how enzyme reaction 
rates depend on substrate concentration and how the key parameters 
V<sub>max</sub> and K<sub>m</sub> influence the response. Altering 
V<sub>max</sub> modifies the maximum rate the enzyme can achieve, while 
adjusting K<sub>m</sub> affects how efficiently the enzyme binds to its 
substrate.
